Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-04-02 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Added text Chapter 3 for graphs.

The natural gas burner provided the most repeatable results, with an average steady state heat release rate of 80~kW~$\pm$~9~kW with an average total heat release of 23.3~MJ~$\pm$~1.6~MJ over a 300~s period.  The gasoline, based on the average over a 30 s interval bounding the peak heat release rate of each of the 17 experiments, provided an average peak heat release rate of 80~kW~$\pm$~24~kW.  The total heat release of the gasoline over the 300~s period after ignition was 13.6~MJ~$\pm$~0.6~MJ.  The polyurethane foam demonstrated the largest variability of the three fuels in terms of peak heat release rate and time to peak heat release.  Averaging over a 15~s interval bounding the peak heat release rate from each of the ten experiments yielded an average peak heat release rate of 45~kW~$\pm$~21~kW.  The total heat release over the 300~s period after ignition was 4.1~MJ~$\pm$~0.2~MJ

@@

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix X.  For the natural gas fueled experiments only two replicates were conducted at each position given the higher level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below,a full set of the averaged data from the natural gas fueled experiments will be presented to demonstrate the impact on distance from the wall on the temperatures in the plume as well as the conditions measured on the wall. Then composite graphs of the wall temperatures and heat flux measure on the wall versus distance will be shown for each fuel type.

@@

Figure~\ref{NCTW_Surf_Center_comp} shows the temperatures of the array of thermocouples located along the vertical centerline of the burners for each of the three fuels.  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average peak temperature of the replicate experiments, for each thermocouple position is shown. Moving from 0.2~m above the top surface of the burner up to 2.2~m above the top of the burner surface in increments of 0.2~m. As in the data analysis for the source fires, the peak averages were based on a 30~s period bounding the peak for the natural gas and gasoline fires.  A 15~s interval bounding the peak was used for the polyurethane, in order to be more representative of the peak conditions.

@@

Figure~\ref{NCTW_HF_Center_Comp} shows the results from the array of total heat flux gauges located along the vertical centerline of the burners for each fuel.  The sensing surface of the gauges were installed flush with the wall surface and exposed to the fire. In each graph, the averaged peak heat flux for each of the replicate experiments and each position is shown.  Moving from 0.2~m above the top surface of the burner up to 1.2~m above the top of the burner surface in increments of 0.2~m.

>>>>

The natural gas burner provided the most repeatable results, with an average steady state heat release rate of 80~kW~$\pm$~9~kW with an average total heat release of 23.3~MJ~$\pm$~1.6~MJ or 11~\% over a 100~s period. It should be noted that 11~\% is also the (2$\sigma$) uncertainty estimate of the calorimeter.  The gasoline, based on the average over a 30 s interval bounding the peak heat release rate of each of the 18 experiments, provided an average peak heat release rate of 80~kW~$\pm$~24~kW or 20~\%.  The total heat release of the gasoline over the 300~s period after ignition was 13.6~MJ~$\pm$~0.6~MJ.  The polyurethane foam demonstrated the largest variability of the three fuels in terms of peak heat release rate and time to peak heat release.  Averaging over a 15~s interval bounding the peak heat release rate from each of the ten experiments yielded an average peak heat release rate of 45~kW~$\pm$~21~kW or 34~\%.  The total heat release over the 300~s period after ignition was 4.1~MJ~$\pm$~0.2~MJ

@@

The results are presented for each of the four burner positions, (0.61~m, 0.31~m, 0.15~m and 0.0~m), and each of the three fuels in Appendix B.  For the natural gas fueled experiments only two replicates were conducted at each position given the higher level of control and repeatability.  For the gasoline and polyurethane fueled experiments, three replicates were conducted at each position. The position is measured as the distance between the edge of the burner closest to the wall and the wall.  The results of the wall temperatures and heat fluxes are presented as averages of the replicate experiments.  In the sections below,a full set of the averaged data from the natural gas fueled experiments will be presented to demonstrate the impact on distance from the wall on the temperatures in the plume as well as the conditions measured on the wall. Then composite graphs of the wall temperatures and heat flux measure on the wall versus distance will be shown for each fuel type.

@@

Figure~\ref{NCTW_Surf_Center_comp} shows the temperatures of the array of thermocouples located along the vertical centerline of the burners for each of the three fuels. Reviewing the graph for each fuel demonstrates how the temperature at the wall decreased as the position of the burner was moved away from the wall. The figures  The beads of the thermocouples were installed in the wall and were flush with the wall surface. In each graph, the average peak temperature of the replicate experiments, for each thermocouple position is shown. Moving from 0.2~m above the top surface of the burner up to 2.2~m above the top of the burner surface in increments of 0.2~m. As with the data analysis for the source fires, the peak averages were based on a 30~s period bounding the peak for the natural gas and gasoline fires.  A 15~s interval bounding the peak was used for the polyurethane, in order to be more representative of the peak conditions.

Figure~\ref{NCTW_Surf_Edge_comp} presents the temperature of the vertical array of thermocouples positioned on the wall along the edge of the burner for each of the three fuels.  The data is treated in the same manner as in Figure~\ref{NCTW_Surf_Center_comp}.  For each fuel the average peak temperatures at the burner edge are significantly lower when the centerline temperatures when the burner is positioned against the wall.  As the burners were moved away from the wall the temperatures at the centerline and the edge thermocouple arrays begin to converge.

Temperatures were also measured on the backside of the wall, along the vertical centerline of the burners, Figure~\{NCTW_Back_Cent_comp}. The temperature difference between the peak average temperatures on the front (fire side) of the wall at 0.2~m above the burner vs the corresponding temperatures on the backside of the wall have a difference of at least 400~$^\circ$C.  It is consistent with the temperature data shown in Figures~\ref{NCTW_Surf_Center_comp} and \ref{NCTW_Surf_Edge_comp}, that as the burners were moved away from the wall the temperature differences are reduced.

@@

Figures~\ref{NCTW_HF_Center_Comp} and {NCTW_HF_Edge_Comp} show the results from the arrays of total heat flux gauges located along the vertical centerline and the edge of the burners for each fuel.  The sensing surface of the gauges were installed flush with the wall surface and exposed to the fire. In each graph, the averaged peak heat flux for each of the replicate experiments and each position is shown.  Moving from 0.2~m above the top surface of the burner up to 1.2~m above the top of the burner surface in increments of 0.2~m.  Reviewing the graph for each fuel demonstrates how the heat flux at the wall decreased as the position of the burner was moved away from the wall.  (Review video and tie in ST TC graph with edge peak NG at 0.15)
